---
title: "Advanced Meta-Analysis Methods"
author: "Mahmood Ahmad"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Meta-Analysis Methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Introduction

This vignette covers advanced meta-analysis techniques available in cbamm:

1. Network Meta-Analysis
2. Bayesian Meta-Analysis
3. Individual Participant Data (IPD) Meta-Analysis
4. Dose-Response Meta-Analysis
5. Machine Learning Methods
6. Multivariate Meta-Analysis

## 1. Network Meta-Analysis

Network meta-analysis (NMA) allows simultaneous comparison of multiple interventions.

### Basic NMA

```{r eval=FALSE}
# Prepare network data
# Each row is a treatment arm within a study
nma_data <- data.frame(
  study = c(1, 1, 2, 2, 3, 3, 4, 4, 4),
  treatment = c("A", "B", "A", "C", "B", "C", "A", "B", "C"),
  effect = c(0, 0.5, 0, 0.8, 0, 0.3, 0, 0.6, 0.7),
  se = c(0.1, 0.15, 0.1, 0.2, 0.12, 0.18, 0.15, 0.18, 0.20),
  baseline = c("A", "A", "A", "A", "B", "B", "A", "A", "A")
)

# Run network meta-analysis
nma_result <- network_meta_analysis(nma_data, reference = "A")
print(nma_result)

# Visualize network structure
plot_network(nma_result)

# Forest plot of all comparisons vs reference
forest_plot_nma(nma_result)

# Rank treatments
rankings <- rank_treatments(nma_result, higher_better = TRUE)
print(rankings)

# Test for inconsistency
inconsistency <- test_inconsistency(nma_result, nma_data)
print(inconsistency)
```

## 2. Bayesian Meta-Analysis

Bayesian methods provide full posterior distributions and natural interpretation.

### Basic Bayesian MA

```{r eval=FALSE}
# Bayesian random-effects meta-analysis
bayes_result <- bayesian_ma(
  data,
  n_iter = 10000,
  n_burn = 2000,
  thin = 5,
  prior_mean = 0,
  prior_sd = 10
)

print(bayes_result)

# Plot posterior distributions
plot(bayes_result)

# Calculate Bayes factor
bf <- bayes_factor_ma(bayes_result, null_value = 0)
print(bf)
```

### Bayesian Meta-Regression

```{r eval=FALSE}
data$year <- 2010:2019

bayes_metareg <- bayesian_metareg(
  data,
  formula = ~ year,
  n_iter = 10000
)

print(bayes_metareg)
```

### Posterior Predictive Checks

```{r eval=FALSE}
# Check model fit
ppc_result <- posterior_predictive_check(bayes_result)
plot(ppc_result)
```

## 3. Individual Participant Data (IPD) Meta-Analysis

IPD meta-analysis uses individual-level data rather than aggregate summaries.

### One-Stage IPD

```{r eval=FALSE}
# IPD data: one row per participant
ipd_data <- data.frame(
  study = rep(1:5, each = 100),
  participant_id = 1:500,
  outcome = rnorm(500, 50, 10),
  treatment = rep(c(0, 1), 250),
  age = rnorm(500, 45, 12),
  sex = sample(c("M", "F"), 500, replace = TRUE)
)

# One-stage analysis
ipd_one <- ipd_onestage(
  data = ipd_data,
  outcome_var = "outcome",
  treatment_var = "treatment",
  covariates = c("age", "sex")
)

print(ipd_one)
```

### Two-Stage IPD

```{r eval=FALSE}
# Two-stage: analyze each study, then pool
ipd_two <- ipd_twostage(
  data = ipd_data,
  outcome_var = "outcome",
  treatment_var = "treatment",
  covariates = c("age", "sex")
)

print(ipd_two)
```

### Treatment-Covariate Interactions

```{r eval=FALSE}
# Test for effect modification
interaction <- ipd_interaction_analysis(
  data = ipd_data,
  outcome_var = "outcome",
  treatment_var = "treatment",
  moderator = "age"
)

print(interaction)
```

## 4. Dose-Response Meta-Analysis

Model the relationship between dose and effect.

### Linear and Non-Linear Models

```{r eval=FALSE}
# Dose-response data
dr_data <- data.frame(
  study = rep(1:10, each = 3),
  dose = rep(c(0, 10, 20), 10),
  effect = c(...),  # Effect at each dose level
  se = c(...)
)

# Linear model
dr_linear <- dose_response_ma(dr_data, model = "linear")
print(dr_linear)

# Quadratic model
dr_quad <- dose_response_ma(dr_data, model = "quadratic")

# Restricted cubic spline
dr_spline <- dose_response_ma(dr_data, model = "restricted_spline", knots = 3)

# Compare models
plot_dose_response(dr_linear)
plot_dose_response(dr_spline)

# Test for non-linearity
test_nonlinearity(dr_spline, dr_data)

# Calculate ED50
ed50 <- calculate_ed50(dr_spline)
print(ed50)
```

## 5. Machine Learning Methods

### Random Forest Meta-Analysis

```{r eval=FALSE}
# Add multiple study characteristics
data$year <- 2010:2019
data$sample_size <- sample(50:500, 10)
data$quality <- sample(1:10, 10, replace = TRUE)
data$country <- sample(c("USA", "UK", "EU"), 10, replace = TRUE)

# Random forest to model heterogeneity
rf_result <- rf_meta_analysis(
  data,
  predictors = c("year", "sample_size", "quality"),
  n_trees = 500
)

print(rf_result)

# Variable importance
print(rf_result$importance)
print(rf_result$heterogeneity_explained)
```

### Neural Network Meta-Analysis

```{r eval=FALSE}
# Neural network for complex non-linear relationships
nn_result <- nn_meta_analysis(
  data,
  predictors = c("year", "quality"),
  hidden_units = 5,
  n_epochs = 1000
)

print(nn_result)
print(nn_result$r_squared)
```

### Automated Outlier Detection

```{r eval=FALSE}
# Isolation forest for outlier detection
outliers <- detect_outliers_ml(data, contamination = 0.1)
print(outliers)

# Remove outliers and reanalyze
data_clean <- data[!outliers$outliers, ]
result_clean <- cbamm_fast(data_clean)
```

### Ensemble Methods

```{r eval=FALSE}
# Combine multiple methods
ensemble_result <- ensemble_meta_analysis(
  data,
  methods = c("DL", "REML", "PM", "ML")
)

print(ensemble_result)
```

## 6. Multivariate Meta-Analysis

Jointly analyze multiple correlated outcomes.

### Basic Multivariate MA

```{r eval=FALSE}
# Data with multiple outcomes per study
mv_data <- data.frame(
  study = rep(1:10, each = 2),
  outcome = rep(c("outcome1", "outcome2"), 10),
  effect = rnorm(20, 0.5, 0.2),
  se = runif(20, 0.1, 0.3)
)

# Multivariate meta-analysis
mv_result <- multivariate_ma(
  mv_data,
  correlation = 0.5  # Within-study correlation
)

print(mv_result)
plot_mvma(mv_result)

# Test difference between outcomes
diff_test <- test_outcome_difference(mv_result, "outcome1", "outcome2")
print(diff_test)
```

### Multivariate Meta-Regression

```{r eval=FALSE}
mv_data$year <- rep(2010:2019, each = 2)

mv_metareg <- multivariate_metareg(
  mv_data,
  formula = ~ year,
  correlation = 0.5
)

print(mv_metareg)
```

## Performance Optimization

For large meta-analyses, use parallel processing:

```{r eval=FALSE}
# Parallel bootstrap
parallel_boot <- parallel_bootstrap_ma(
  data,
  n_bootstrap = 10000,
  n_cores = 4
)

# Parallel subgroup analysis
data$subgroup <- sample(c("A", "B", "C"), nrow(data), replace = TRUE)
parallel_subgroups <- parallel_subgroup_analysis(
  data,
  subgroup_var = "subgroup",
  n_cores = 4
)

# Batch processing multiple meta-analyses
data_list <- list(data1, data2, data3, ...)
batch_results <- batch_meta_analysis(
  data_list,
  parallel = TRUE,
  n_cores = 4
)
```

## Integration with Other Packages

```{r eval=FALSE}
# Convert to metafor format
metafor_data <- to_metafor(data)

# Use metafor's advanced features
if (require(metafor)) {
  rma_result <- metafor::rma(yi = yi, sei = sei, data = metafor_data)
}

# Convert back to cbamm
cbamm_result <- from_metafor(rma_result)

# Use metafor diagnostics
diagnostics <- metafor_diagnostics(
  data,
  diagnostics = c("influence", "gosh", "trimfill")
)
```

## Choosing the Right Method

```{r eval=FALSE}
# Auto-select best method based on data
recommendation <- auto_select_method(data)
print(recommendation)

# Compare multiple methods
comparison <- compare_methods(
  data,
  methods = c("DL", "REML", "PM", "ML")
)
print(comparison)
```

## Conclusion

cbamm provides state-of-the-art meta-analysis methods. Choose the appropriate
technique based on your research question:

- **Multiple treatments?** → Network meta-analysis
- **Complex moderator relationships?** → ML methods or Bayesian models
- **Individual-level data?** → IPD meta-analysis
- **Dose-response relationship?** → Dose-response models
- **Multiple outcomes?** → Multivariate meta-analysis
- **Continuous updating?** → See `vignette("living_reviews")`
