---
title: "Getting Started with cbamm"
author: "Mahmood Ahmad"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with cbamm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

The **cbamm** package provides comprehensive tools for conducting meta-analyses,
from basic pooling to advanced techniques like machine learning-based methods,
living systematic reviews, and multivariate meta-analysis.

## Installation

```{r eval=FALSE}
# Install from GitHub
devtools::install_github("mahmood726-cyber/LFA")

# Load the package
library(cbamm)
```

## Basic Meta-Analysis

### Preparing Your Data

cbamm requires a data frame with at least three columns:
- `study`: Study identifier
- `effect`: Effect size estimate (e.g., SMD, log OR, log RR)
- `se`: Standard error of the effect size

```{r eval=FALSE}
# Example data
data <- data.frame(
  study = paste0("Study ", 1:10),
  effect = c(0.5, 0.3, 0.7, 0.4, 0.6, 0.5, 0.2, 0.8, 0.4, 0.5),
  se = c(0.15, 0.20, 0.18, 0.16, 0.14, 0.17, 0.22, 0.15, 0.19, 0.16)
)
```

### Running a Basic Meta-Analysis

```{r eval=FALSE}
# Fast random-effects meta-analysis
result <- cbamm_fast(data, method = "DL")
print(result)

# The result contains:
# - estimate: Pooled effect size
# - ci_lower, ci_upper: 95% confidence interval
# - tau2: Between-study variance
# - I2: Heterogeneity statistic (%)
# - Q: Cochran's Q statistic
# - p_value: P-value for pooled estimate
```

### Visualizing Results

```{r eval=FALSE}
# Forest plot (base graphics)
plot(result)

# Modern ggplot2 forest plot
library(ggplot2)
gg_forest_plot(data, result = result, show_weights = TRUE)

# Funnel plot for publication bias
gg_funnel_plot(data, result = result, show_contours = TRUE)
```

## Assessing Publication Bias

```{r eval=FALSE}
# Egger's test
egger_result <- egger_test(data)
print(egger_result)

# PET-PEESE correction
pet_peese_result <- pet_peese(data, method = "conditional")
print(pet_peese_result)

# Trim-and-fill
tf_result <- trim_fill(data)
print(tf_result)
```

## Sensitivity Analysis

```{r eval=FALSE}
# Leave-one-out analysis
loo_result <- leave_one_out(data)
print(loo_result)

# Influence diagnostics
influence_result <- influence_diagnostics(data)
print(influence_result)

# Baujat plot to identify outliers
baujat_plot(data)
```

## Meta-Regression

When you have study-level characteristics (moderators):

```{r eval=FALSE}
# Add moderator variables
data$year <- 2010:2019
data$quality <- sample(1:10, 10, replace = TRUE)

# Meta-regression
metareg_result <- meta_regression(data, formula = ~ year + quality)
print(metareg_result)

# Visualize relationship
gg_metareg_plot(data, moderator = "year", show_bubble = TRUE)
```

## Subgroup Analysis

```{r eval=FALSE}
# Add subgroup variable
data$intervention_type <- rep(c("Drug A", "Drug B"), 5)

# Subgroup meta-analysis
subgroup_result <- subgroup_analysis(data, subgroup_var = "intervention_type")
print(subgroup_result)
```

## Cumulative Meta-Analysis

Track how evidence accumulates over time:

```{r eval=FALSE}
# Cumulative meta-analysis ordered by year
cumulative_result <- cumulative_meta_analysis(data, order_by = "year")
print(cumulative_result)

# Plot evolution of evidence
gg_cumulative_plot(cumulative_result)
```

## Effect Size Conversions

cbamm provides 17 functions for converting between effect size metrics:

```{r eval=FALSE}
# Convert Cohen's d to Hedges' g
g <- d_to_g(d = 0.5, n1 = 50, n2 = 50)

# Convert odds ratio to risk ratio (given baseline risk)
rr <- or_to_rr(or = 2.0, baseline_risk = 0.3)

# Convert correlation to Cohen's d
d <- r_to_d(r = 0.3)

# Fisher z-transformation for correlations
z <- fisher_z(r = 0.3)
```

## Prediction Intervals

Go beyond confidence intervals to predict effects in new studies:

```{r eval=FALSE}
# Calculate prediction interval
pi_result <- prediction_interval(data, method = "REML")
print(pi_result)

# Prediction intervals account for heterogeneity
# CI tells you about the average effect
# PI tells you the range for a new study
```

## Next Steps

- See `vignette("advanced_methods")` for network meta-analysis, Bayesian methods, and ML approaches
- See `vignette("living_reviews")` for maintaining living systematic reviews
- See `vignette("multivariate")` for analyzing multiple correlated outcomes
- See `vignette("diagnostic_accuracy")` for diagnostic test accuracy meta-analysis

## Quick Reference

| Task | Function |
|------|----------|
| Basic meta-analysis | `cbamm_fast()` |
| Publication bias | `egger_test()`, `pet_peese()`, `trim_fill()` |
| Sensitivity analysis | `leave_one_out()`, `influence_diagnostics()` |
| Meta-regression | `meta_regression()` |
| Subgroup analysis | `subgroup_analysis()` |
| Cumulative MA | `cumulative_meta_analysis()` |
| Prediction intervals | `prediction_interval()` |
| Forest plots | `gg_forest_plot()` |
| Funnel plots | `gg_funnel_plot()` |
| Interactive dashboard | `launch_meta_dashboard()` |

## Getting Help

```{r eval=FALSE}
# Function documentation
?cbamm_fast
?gg_forest_plot

# Package documentation
help(package = "cbamm")

# Report issues
# https://github.com/mahmood726-cyber/LFA/issues
```
